Open Preprints in Ecology \& Evolution

### 2012-09-15 ms.tex
added some edits to the intro, notes (in comments) about further discussion points, and added some comments to the discussion.

\keywords{Publishing; arXiv; Green Open Access.}

@@

Open preprints servers allow authors to make their manuscripts publicly
available before, or in parallel to, submitting them to journals for traditional
peer-review. The idea became popular with arXiv, an open preprint server, that
started 20 years ago \cite{gin11}. The idea is fundamentally simple: to make the
science available to the entire scientific community as soon as its done instead
of having to wait for the peer-review process. The point of arXiv and open
preprints servers is not to avoid the peer-review process. Almost all
manuscripts on arXiv are submitted to peer-review. The point is to open an
important phase of the publication cycle ...

@@

There are other ``cultural'' barriers to adoption of preprint servers
by ecologists and evolutionary biologists.  The most notable preprint
server, arXiv, is designed to work best with \TeX{}, a nearly
universal document preparation system among physicists and
mathematicians.  Jackson \cite{jackson2002preprints} further argues
that \TeX{} introduced an ``open source mindset'' to its users, who
now freely share their research findings as well as their software.
Many biologists instead use proprietary software to prepare their
research findings, and many ecology and evolution journals officially
prefer Microsoft Word documents as submissions.  The recent
discipline-wide adoption of free software packages such as R, as well
as interest in open access publishing, has coincided with recent use
of preprint servers by biologists, supporting Jackson's claims.

@@

Posting preprints online increases the community of available informal
peer reviewers, and uses the internet for its original
community-building purposes.  Preprint servers also facilitate
communication between disciplines, bridging cultural as well as
geographic divides.
% JA: perhaps something less advertising-like should go here:
The advantages are clear and the costs are low.
% BUY NOW!

>>>>

\author{Karthik Ram}
\affiliation{Environmental Science, Policy, and Management. University of California, Berkeley. Berkeley, CA. United-States of America.}

\keywords{Publishing; arXiv; pre print server, Green Open Access.}

@@

Open preprints servers allow authors to make their manuscripts publicly available before, or in parallel to, submitting them to journals for traditional peer-review. This idea gained popularity 20 years ago with the advent of arXiv, an open preprint server, widely used in the physical sciences \cite{gin11}. The idea behind pre-print servers is fundamentally simple: to make the results of a scientific endavour available to the scientific community as soon as possible rather than wait until the peer-review process is fully completed. The point of arXiv and open preprints servers is not to avoid the peer-review process. Almost all manuscripts on arXiv are submitted to peer-review.  Sharing drafts of research papers on pre-print servers has numerous advantages. These include rapid dissemination of work in progress to a wider audience that includes not only scientists and the general public, but also readers in developing countries where access to subscription journals are often a limiting factor. Further, pre-prints provide the opportunity to solicit feedback from a larger pool of reviewers. As a consequence, pre print servers increase the number of opportunities for review and revision prior to publication, resulting in higher quality submissions which could also alleviate reviewer burden.

@@

There are other ``cultural'' barriers to adoption of preprint servers
by ecologists and evolutionary biologists.  The most notable preprint
server, arXiv, is designed to work best with \TeX{}, a nearly
universal document preparation system among physicists and
mathematicians.  Jackson \cite{jackson2002preprints} further argues
that \TeX{} introduced an ``open source mindset'' to its users, who
now freely share their research findings as well as their software.
Many biologists instead use proprietary software to prepare their
research findings, and many ecology and evolution journals officially
prefer Microsoft Word documents as submissions.  The recent
discipline-wide adoption of free software packages such as R and associated  document generation systems \cite{xie2012}, as well
as interest in open access publishing, has coincided with recent rise in use
of preprint servers by biologists, supporting Jackson's claims.
% Please add this to your citation list: Xie Y (2012). knitr: A general-purpose pack­age for dynamic report gen­er­a­tion in R. R pack­age ver­sion 0.4.1, .

@@

Posting preprints online increases the community of available informal
peer reviewers, and uses the internet for its original
community-building purposes.  Preprint servers also facilitate
communication between disciplines, bridging cultural as well as
geographic divides.
% JA: perhaps something less advertising-like should go here:
The advantages are clear and the costs are low.

% Karthik's comments: 
% It would be worthwhile to mention that not everyone is stoked about pre-print servers (crazy people). So maybe cite something to that effect?
% ACS Journal Editors’ Policy on Preprints at http://pubs.acs.org/cen/ preprint.% html. Criticisms include: 
% flooding the web with too many repetitive pubs
% premature results without any support from data
% etc.

% BUY NOW!
